Testing Quasar Unification: I. Radiative Transfer in Clumpy Winds

### 2015-12-17 paper_mnras.tex
about to clean up

Using this prescription, we conducted a limited parameter
search over a 5-dimensional parameter space involving the 
variables $r_{min}$, $\theta_{min}$, $f_V$, $\alpha$ and $R_v$.
The grid points are shown in table 1.
The aim here was to first fix $M_{BH}$ and $\dot{m}$ to their H13 values,
and increase $L_X$ to $10^{45}$~erg~s$^{-1}$ (a more realistic value for a 
quasar of $10^9M_\odot$ and an eddington fraction of $0.2$; see section~\ref{sec:xray}).

We then evaluated these models based on whether their synthetic spectra exhibited the 
following properties of quasars and BALQSOs:

\begin{itemize}
\item UV absorption lines 
with $BI > 0$ at $\sim20\%$ of viewing angles (e.g. Knigge et al. 2010)
\item Line emission emerging at low inclinations, with $EW\sim40\AA$ in \civ\ \citep[e.g. ][]{shen2011}
\item H recombination lines with $EW\sim50\AA$ in \la\ \citep[e.g. ][]{shen2011}
\item  \mg\ and \al\ (LoBAL) absorption features with $BI > 0$ at a small subset of BAL angles
\item Verisimilitude with quasar composite spectra.
\end{itemize}
Where $BI$ is the `Balnicity Index' (Weymann et al. 1991), given by
\begin{equation}
BI = \int^{25000}_{3000} \left( 1 - \frac{f(v)}{0.9} \right) dv.
\end{equation}
The constant $C=0$ everywhere, unless the normalized flux
has satisfied $f(v)<0.9$ continuously for at least $2000$~km~s$^{−1}$, 
whereby $C$ is set to $1$.

@@

The total line luminosity also increases dramatically compared to H13. 
This is due to heating of the plasma by the increased X-ray 
luminosity, which avoids over-ionization due to clumping.
This line emission typically emerges on the edge of the wind
nearest the central source. The location of the line emitting regions
is dependent on the ionization state, as well as the incident X-rays.
The radii of these emitting regions is important,
and can be compared to observations. The line luminosities, $L$,
shown in the figure correspond to the luminosity in $\LUM$ of photons
escaping the Sobolev region for each line. 
% This is equivalent to $\beta_{ul} n_u A_{ul}$, where the three quantities
% represent the Sobolev escape probability, upper level number density and 
% Einstein A coefficient for the line. 
As shown in figure~\ref{fig:wind},
the \civ\ line in our model is typically formed between 
$100-1000~r_G$ ($\sim10^{17}-10^{18}~\rm{cm}$).
This is in rough agreement with the reverberation mapping 
results of Kaspi (2000) for the $2.6\times10^{9} M_\odot$ quasar S5 0836+71,
and also compares favourably with microlensing measurements of the size of the
\civ\ emission line region in the BALQSO H1413+117 \citep{odowd2015}.

@@

\begin{figure*}
\centering
\includegraphics[width=1.0\textwidth]{figures/uvspec_new.png}
\caption
{
Synthetic spectra at four viewing angles in our model. At 
$20^\circ$ and $60^\circ$ we show a comparison to an SDSS quasar composite
from Recihard et al. (2003). At $73^\circ$ and $76^\circ$ we show a comparison to
an {\sl HST} STIS spectrum of the high BALnicity BALQSO 
PG0946+301 (Arav et al. 2000), and an SDSS spectrum of the LoBAL quasar 
XXXXX (REF), respectively. The dotted line shows a disc
only continuum to show the effect of the outflow on the continuum level. 
All the spectra are scaled to the model flux at $2000$\AA, expect for the 
{\sl HST} STIS spectrum of PG0946+301, which is scaled using a continuum fit
due to the incomplete wavelength coverage.
% HST STIS spectrum of PG0946+301 (Arav et al. 2000)
}
\label{fig:uvspec}
\end{figure*}

\begin{figure} 
\centering
\includegraphics[width=0.5\textwidth]{figures/windnew3.png}
\caption
{
A cartoon describing the broad classes of sightline 
in our model, illustrating how geometric effects lead to 
the different emergent spectra. The colour gradient is approximate,
but indicates the stratified ionization structure, 
from highly ionized (yellow) to low ionization (purple) material.
%[FIGURE NEEDS IMPROVING]
}
\label{fig:sightline}
\end{figure}

\noindent
Figure~\ref{fig:uvspec} shows the synthetic spectrum in the UV from the fiducial model. 
To assess the ability of the synthetic spectra to match real 
quasar spectra, we also show {\sl Sloan Digital Sky Survey} (SDSS) quasar
composites from \cite{reichard2003}, normalised to the flux at 2000\AA\
for low inclinations. Unfortunately, the wide variety of
line profile shapes and internal trough structure in BALQSOs
tends to `wash out' BAL troughs in composite spectra
to the extent that BALQSO composites do not resemble typical BALQSOs.
Because of this, we instead compare to a {\sl Hubble Space Telescope} 
STIS spectrum of the high BALnicity BALQSO PG0946+301 (Arav et al. 2000),
and an SDSS spectrum of the LoBAL quasar XXXXX (REF),
for the angles of $73^\circ$ and $76^\circ$, respectively. 
We show a cartoon illustrating how geometric effects determine
the output spectra in figure~\ref{fig:sightline}.

@@

Non-black saturation is observed in the absorption troughs of BALQSOs \citep{arav1999b,arav1999}.
This is usually explained either as  partial covering of the continuum
source or by scattered contributions to the BAL troughs, necessarily
from an opacity source not co-spatial with the BAL forming region.
The scattered light explanation is supported by spectropolarimetry results
\citep{lamy2000}. Our spectra do not show nonblack saturation.
Instead, we find black, saturated troughs at angles $i > 73^\circ$, and the BALs
are non-saturated at lower inclinations. The reasons for this are inherent in the construction of our model. 
First, the microclumping assumption does not allow for 
porosity in the wind, meaning that it does not naturally produce
a partial covering absorber. To allow this, an alternative approach
such as {\em macroclumping} would be required \citep[e.g.][]{hamann2008,surlan2012}.
Second, our wind does not have a significant scattering 
contribution along sightlines which do not pass through the BAL region,
meaning that any scattered component to the BAL troughs is absorbed by line opacity.
This suggests that either the scattering cross-section of the wind must
be increased (with higher mass loss rates or covering factors), or 
that an additional source of electron opacity is required, potentially
in a polar direction above the disc.  
% {\bf KSL: it sounds like what you are suggesting is a low density inner flow, our domain 2 stuff actually permits ... or indeed one might be able to produce this within SV is you adopted a different mass loss with radius assumption. JM: I am being deliberately 
% vague about what I am suggesting. I am saying there are three physical
% possibilities: either wind itself causes this scattering contribution, which doesn't happen in our model, or you have partial covering, or you have some additional scatterer.}

@@

We find significant collisionally excited line emission emerges
at low inclinations in the synthetic spectra, particular in the \civ\ line.
The improved treatment of recombination also results in a strong \la\ line. 
In the context of unification, this is a promising result, 
and shows that a biconical wind can produce significant 
emission at `quasar-like' angles. 
The spectra do not contain the strong C~\textsc{iii}]~1909\AA\, 
line seen in the quasar composite spectra. 
This is because we do not yet treat C as a full macro-atom with a 
full collisional rates between forbidden
or semi-forbidden transitions, as would be required.
The critical density of the C~\textsc{iii}]~1909\AA\, line 
is $n_e\sim10^{9.5}~\rm{cm^{-3}}$ \citep{wei1988}, which is higher than much of the 
outer portion of our wind. 
We therefore expect a model with these 
parameters to produce a C~\textsc{iii}]~1909\AA\ line
with a proper treament.

The model produces strong emission lines in \civ, \nv\ and \la,
as well as a weak \mg\ line. The shapes and widths of these lines
match the composites fairly well. However, the line-to-continuum ratios 
at low inclinations in our model are significantly weaker than the quasar 
composites. Increasing the density of the outflow, by altering the mass 
loss rate  or velocity law, can produce more line emission.
However, the red wing of the BAL profiles is generally stronger than 
seen in BALQSO spectra and composites. This illustrates a fundamental 
problem with a geometric unification model such as this:
that the line-to-continuum ratios at high inclinations are
significantly affected by disc foreshortening and limb darkening.
The angular distribution of the disc radiation is clearly
crucially important in determining the emergent line ratios.  
% {\bf KSL: This is a bit oddly explained. Our problem is at low inclination.  Note that I am a bit concerned with Fig 3; the model spectra and the observations look very different.}

@@

\begin{figure*}
\centering
\includegraphics[width=1.0\textwidth]{figures/lx_a05_pre.png}
\caption
{
X-ray ($2$~keV) luminosity of the our clumped model (red squares) 
and the H13 model (blue squares), plotted against monochromatic luminosity 
at 2500\AA. The points are labeled according to inclination; angles
$>70^\circ$ correspond to BALs in our scheme (see figure 4).
Also plotted are masurements from 
the COMBI-7 AGN and the BQS samples (Steffen et al. 2006) and the Saez et al. (2012) 
sample of BALQSOs. The dotted line shows the best fit relation for non-BALQSOs 
from Steffen et al. (2006).
}
\label{fig:xray}
\end{figure*}

@@

The low inclination, `BALQSO-like' viewing angles show moderate agreement with the data,
and are X-ray weak due to bound-free and electron scattering opacities in the wind.
Typically, BALQSOs show strong X-ray absorption with columns 
of $N_H\sim10^{23}~\rm{cm^{-2}}$ 
\citep{green1996,mathur2000,green2001,grupemathur2003}.
This is often cited as evidence that the BAL outflow is shielded from
the X-ray source, especially as sources with strong X-ray absorption tend
to exhibit deep BAL troughs and high outflow velocities 
\citep{brandt2000,laorbrandt2002,gallagher2006}.
Our results imply that the clumpy BAL outflow
itself can be responsible for the strong X-ray absorption, 
and supports Hamann et al.'s (2013) suggestion that 
this explains the weaker X-ray absorption in mini-BALs 
compared to BALQSOs.

@@

\begin{figure*}
\centering
\includegraphics[width=1.0\textwidth]{figures/c4_angles.png}
\caption
{
\civ , \mg , \al\ and Fe~\textsc{ii} line profiles for viewing angles
from $72-79^\circ$. The profiles are plotted relative to the local
continuum with an offset applied for clarity. Lower ionization
profiles appear at a subset of high inclinations, compared
to the ubiquitous \civ\ profile.
}
\label{fig:lobal}
\end{figure*}

At certain sightlines, the synthetic spectra exhibit blue-shifted BALs in \al\ and \mg --
the absorption lines seen in LoBALQSOs, and we even see absorption in Fe~\textsc{ii}
at the highest inclinations. Line profiles in velocity space 
for \civ, \al\ and \mg, are shown in figure~\ref{fig:lobal} for a range
of BALQSO viewing angles. We find that ionization stratification
of the wind causes lower ionization material to have a smaller covering factor, 
as demonstrated by figures~\ref{fig:wind} and \ref{fig:lobal}.
This confirms the behaviour expected from a unification model such as Elvis (2000). 
LoBALs are only present at viewing angles close to edge-on ($i>75^\circ$),
as predicted by polarisation results \citep{brotherton1997}.
As observed in a BALQSO sample by \cite{filizak2014}, we find that
BAL troughs are wider and deeper when low ionization absorption features are present,
and high ionization lines have higher blue-edge velocities than the 
low ionization species.
There is also a correlation between the strength of LoBAL features
and the amount of continuum attenuation at that sightline, particularly
blueward of the Lyman edge as the low ionization base 
intersects the line-of-sight. 
A model such as this therefore predicts that LoBALQSOs and FeLoBALQSOs 
have stronger Lyman edge absorption and 
are more Compton-thick than HiBALQSOs and Type 1 quasars.
An edge-on scenario also offers a potential explanation for the rarity of LoBAL and
FeLoBAL quasars, due to a foreshortened and attenuated continuum, 
although, as noted in section~\ref{angular}, BAL fraction 
inferences are fraught with complex selection effects.

@@

 \section{Discussion: Parameter Sensitivity and Inclination Dependence}

 We have demonstrated that our fiducial model successfully reproduces much of
 the phenomenology of luminous AGN and quasars, with a few exceptions.
 Here, we discuss these shortcomings, by discussing parameter sensitivity
 and the dependence of line EWs on inclination.

\subsection{The Angular Distribution of Line And Continuum Emission}
\label{sec:angular}

\begin{figure}
\centering
\includegraphics[width=0.49\textwidth]{figures/hist_ew.png}
\caption
{
{\sl Top Panel:} A histogram of $1/2$ equivalent widths ($W_{\lambda}/2$)
for BALQSOs and non-BAL quasars from the {\sl SDSS DR7} quasar sample (Shen et al. 2011).
{\sl Bottom Panel:} $W_{\lambda,RW}$ as a function of inclination in our example model (black line).
The green shaded area indicates viewing angles which look into the wind (`BALQSO-like').
}
\label{fig:ew}
\end{figure}

% \begin{figure}
% \centering
% \includegraphics[width=0.45\textwidth]{figures/ew.png}
% \caption
% {
% $W_{\lambda,RW}$ as a function of inclination in our example model (red line).
% We also the mean $1/2$ equivalent width from DiPompeo et al. (2012) as a horizontal
% line, with the shaded area corresponding to the standard deviation.
% }
% \label{fig:ew}
% \end{figure}

In order to quantitatively assess how emission lines change with 
inclination when blue-shifted absorption 
may affect the line profile, we define the `red wing equivalent width' ($W_{\lambda,RW}$) as
\begin{equation}
W_{\lambda,RW} = \int_{\lambda_0}^{\lambda^\prime} \left( 1 - \frac{F_\lambda}{F_0} \right) d\lambda
\label{rwew}
\end{equation}
where $F_0$ is the continuum flux and the integral is calculated from $\lambda_0$, line centre,
to a wavelength $\lambda^\prime$ where the flux has returned to the continuum level.
This quantity is shown as a function of inclination in figure 5 for the \civ\ and \mg\ UV lines.
We also show the $W_{\lambda,RW}$ expected from isotropic line emission and a foreshortened 
and limb darkened disc  as well as $1/2$ equivalent widths from \cite{dipompeo2012b}.

>>>>

Using this prescription, we conducted a limited parameter
search over a 5-dimensional parameter space involving the 
variables $r_{min}$, $\theta_{min}$, $f_V$, $\alpha$ and $R_v$.
The grid points are shown in table 1.
The aim here was to first fix $M_{BH}$ and $\dot{m}$ to their H13 values,
and increase $L_X$ to $10^{45}$~erg~s$^{-1}$ (a more realistic value for a 
quasar of $10^9M_\odot$ and an Eddington fraction of $0.2$; see section~\ref{sec:xray}).

We then evaluated these models based on 
how closely their synthetic spectra reproduced the 
following properties of quasars and BALQSOs:

\begin{itemize}
\item UV absorption lines 
with $BI > 0$ at $\sim20\%$ of viewing angles (e.g. Knigge et al. 2010)
\item Line emission emerging at low inclinations, with $EW\sim40\AA$ in \civ\ \citep[e.g. ][]{shen2011}
\item H recombination lines with $EW\sim50\AA$ in \la\ \citep[e.g. ][]{shen2011}
\item  \mg\ and \al\ (LoBAL) absorption features with $BI > 0$ at a subset of 
BAL viewing angles
\item Verisimilitude with quasar composite spectra.
\end{itemize}
Here $BI$ is the `Balnicity Index' (Weymann et al. 1991), given by
\begin{equation}
BI = \int^{25000}_{3000} \left( 1 - \frac{f(v)}{0.9} \right) dv.
\end{equation}
The constant $C=0$ everywhere, unless the normalized flux
has satisfied $f(v)<0.9$ continuously for at least $2000$~km~s$^{−1}$, 
whereby $C$ is set to $1$.

@@

The total line luminosity also increases dramatically compared to H13. 
This is primarily due to the higher X-ray luminosity increasing the bound-free
heating in the plasma, which avoids over-ionization due to clumping.
This line emission typically emerges on the edge of the wind
nearest the central source. The location of the line emitting regions
is dependent on the ionization state, as well as the incident X-rays.
The radii of these emitting regions is important,
and can be compared to observations. The line luminosities, $L$,
shown in the figure correspond to the luminosity in $\LUM$ of photons
escaping the Sobolev region for each line. 
% This is equivalent to $\beta_{ul} n_u A_{ul}$, where the three quantities
% represent the Sobolev escape probability, upper level number density and 
% Einstein A coefficient for the line. 
As shown in figure~\ref{fig:wind},
the \civ\ line in the fiducial model is typically formed between 
$100-1000~r_G$ ($\sim10^{17}-10^{18}~\rm{cm}$).
This is in rough agreement with the reverberation mapping 
results of Kaspi (2000) for the $2.6\times10^{9} M_\odot$ quasar S5 0836+71,
and also compares favourably with microlensing measurements of the size of the
\civ\ emission line region in the BALQSO H1413+117 \citep{odowd2015}.

@@

\begin{figure*}
\centering
\includegraphics[width=1.0\textwidth]{figures/uvspec_new.png}
\caption
{
Synthetic spectra at four viewing angles for the fiducial model. At 
$20^\circ$ and $60^\circ$ we show a comparison to an SDSS quasar composite
from Recihard et al. (2003). At $73^\circ$ and $76^\circ$ we show a comparison to
an {\sl HST} STIS spectrum of the high BALnicity BALQSO 
PG0946+301 (Arav et al. 2000), and an SDSS spectrum of the LoBAL quasar 
SDSS J162901.63+453406.0, respectively. The dotted line shows a disc
only continuum to show the effect of the outflow on the continuum level. 
All the spectra are scaled to the model flux at $2000$\AA, expect for the 
{\sl HST} STIS spectrum of PG0946+301, which is scaled using a continuum fit
due to the incomplete wavelength coverage.
% HST STIS spectrum of PG0946+301 (Arav et al. 2000)
}
\label{fig:uvspec}
\end{figure*}

\begin{figure} 
\centering
\includegraphics[width=0.5\textwidth]{figures/windnew3.png}
\caption
{
A cartoon describing the broad classes of sightline 
in the fiducial model, illustrating how geometric effects lead to 
the different emergent spectra. The colour gradient is approximate,
but indicates the stratified ionization structure, 
from highly ionized (yellow) to low ionization (purple) material.
%[FIGURE NEEDS IMPROVING]
}
\label{fig:sightline}
\end{figure}

\noindent
Figure~\ref{fig:uvspec} shows the synthetic spectrum in the UV from the fiducial model. 
To assess the ability of the synthetic spectra to match real 
quasar spectra, we also show {\sl Sloan Digital Sky Survey} (SDSS) quasar
composites from \cite{reichard2003}, normalised to the flux at 2000\AA\
for low inclinations. Unfortunately, the wide variety of
line profile shapes and internal trough structure in BALQSOs
tends to `wash out' BAL troughs in composite spectra
to the extent that BALQSO composites do not resemble typical BALQSOs.
Because of this, we instead compare to a {\sl Hubble Space Telescope} 
STIS spectrum of the high BALnicity BALQSO PG0946+301 (Arav et al. 2000),
and an SDSS spectrum of the LoBAL quasar SDSS J162901.63+453406.0,
for the angles of $73^\circ$ and $76^\circ$, respectively. 
We show a cartoon illustrating how geometric effects determine
the output spectra in figure~\ref{fig:sightline}.

@@

Non-black saturation is observed in the absorption troughs of BALQSOs \citep{arav1999b,arav1999a}.
This is usually explained either as  partial covering of the continuum
source or by scattered contributions to the BAL troughs, necessarily
from an opacity source not co-spatial with the BAL forming region.
The scattered light explanation is supported by spectropolarimetry results
\citep{lamy2000}. Our spectra do not show nonblack saturation.
Instead, we find black, saturated troughs at angles $i > 73^\circ$, and the BALs
are non-saturated at lower inclinations. The reasons for this are inherent in the construction of our model. 
First, the microclumping assumption does not allow for 
porosity in the wind, meaning that it does not naturally produce
a partial covering absorber. To allow this, an alternative approach
such as {\em macroclumping} would be required \citep[e.g.][]{hamann2008,surlan2012}.
Second, our wind does not have a significant scattering 
contribution along sightlines which do not pass through the BAL region,
meaning that any scattered component to the BAL troughs is absorbed by line opacity.
This suggests that either the scattering cross-section of the wind must
be increased (with higher mass loss rates or covering factors), or 
that an additional source of electron opacity is required, potentially
in a polar direction above the disc.  
% {\bf KSL: it sounds like what you are suggesting is a low density inner flow, our domain 2 stuff actually permits ... or indeed one might be able to produce this within SV is you adopted a different mass loss with radius assumption. JM: I am being deliberately 
% vague about what I am suggesting. I am saying there are three physical
% possibilities: either wind itself causes this scattering contribution, which doesn't happen in our model, or you have partial covering, or you have some additional scatterer.}

@@

\begin{table}
\begin{tabular}{p{2cm}p{2cm}p{3cm}}
\hline Property & Synthetic, $20^\circ$ & Observed  (S11) \\ 
\hline \hline
$\log L$[C~\textsc{iv}]  & $44.60$ & $44.42 \pm 0.32$  \\
$\log L$[Mg~\textsc{ii}] & $43.92$ & $43.54 \pm 0.28$  \\
$\log (\nu L_{\nu})_{1350}$  & $46.42$ & $46.01 \pm 0.30$ \\
$\log (\nu L_{\nu})_{3000}$  & $46.18$ & $45.79 \pm 0.30$ \\
% $\log L$[H~$\alpha$] & $43.46$ & $42.86$ & ?? \\
% $EW$[C~\textsc{iv}] (erg~s$^{-1}$) & $3.96 \times 10^{44}$ & $2.62 \times 10^{44}$  \\
% $EW$[Mg~\textsc{ii}] (erg~s$^{-1}$) & $3.50 \times 10^{43}$ & $8.32 \times 10^{43}$ \\
% $EW$[H~$\alpha$] (erg~s$^{-1}$) & $3.50 \times 10^{43}$ & $8.32 \times 10^{43}$ \\
% $\log L_{1350}$ (erg~s$^{-1}$~Hz$^{-1}$) & $46.42$ & $45.56$ & $46.01 \pm 0.30$ \\
% $\log L_{3000}$ (erg~s$^{-1}$~Hz$^{-1}$) & $46.18$ & $45.36$ & $45.79 \pm 0.30$ \\
% $L_{bol})$ (erg~s$^{-1}$) & 2.62
% % $L_{1300})$(C\textsc{iv}) & 
% \\
% \multicolumn{3}{|l|}{At $73^\circ$ (`BALQSO-like')} \\  
% $\log L$[C~\textsc{iv}] (erg~s$^{-1}$) & $44.14$ & ??  \\
% $\log L$[Mg~\textsc{ii}] (erg~s$^{-1}$) & $42.82$ & $44.41$  \\
% $\log L$[H~$\alpha$] (erg~s$^{-1}$) & $42.86$ & ?? \\
% % $EW$[C~\textsc{iv}] (erg~s$^{-1}$) & $3.96 \times 10^{44}$ & $2.62 \times 10^{44}$  \\
% % $EW$[Mg~\textsc{ii}] (erg~s$^{-1}$) & $3.50 \times 10^{43}$ & $8.32 \times 10^{43}$ \\
% % $EW$[H~$\alpha$] (erg~s$^{-1}$) & $3.50 \times 10^{43}$ & $8.32 \times 10^{43}$ \\
% $\log L_{1350}$ (erg~s$^{-1}$~Hz$^{-1}$) & $45.56$ & $46.00$ \\
% $\log L_{3000}$ (erg~s$^{-1}$~Hz$^{-1}$) & $45.36$ & $45.79$ \\
\hline
\end{tabular}
\caption{
Some derived spectral properties of the fiducial model, at $20^\circ$,
compared to observations. The observed values are taken from the Shen et al. (2011)
SDSS DR7 Quasar catalog, and correspond to mean values with standard deviations in log space
from a subsample with $8.5>\log(M_{BH})<9.5$ and $1.5<\log (L_{bol}/L_{edd}) < 0$,
where the BH mass is a \civ\ virial estimate. 
Units are logarithms of values in erg~s$^{-1}$.
%unless stated.otherwise.
}
\label{line_lums}
\end{table}

We find significant collisionally excited line emission emerges
at low inclinations in the synthetic spectra, particular in the \civ\ and \nv\
lines. The improved treatment of recombination also results in a strong \la\ line. 
In the context of unification, this is a promising result, 
and shows that a biconical wind can produce significant 
emission at `quasar-like' angles. To demonstrate this further,
we show line luminosities and monochromatic continuum luminosities
from the synthetic spectra in table~\ref{line_lums}. These are compared to
mean values from a subsample of the SDSS DR7 quasar catalog (Shen et al. 2011) 
with BH mass and Eddington fraction estimates similar to the fiducial model values 
(see caption). The spectra do not contain the strong 
C~\textsc{iii}]~1909\AA\, line seen in the quasar composite spectra,
because semi-forbidden (intercombination)
lines are not included in our modelling.
% This is because we do not fully solve the C level populations,
% and thus the level populations do not accurately reflect those
% as a full macro-atom with a 
% full collisional rates between forbidden
% or semi-forbidden transitions, as would be required.
The critical density of the C~\textsc{iii}]~1909\AA\, line 
is $n_e\sim10^{9.5}~\rm{cm^{-3}}$ \citep{wei1988}, which is higher than much of the 
outer portion of our wind. 
We therefore expect a model with these 
parameters to produce a C~\textsc{iii}]~1909\AA\ line
with a proper treament. We note, however, the the presence of
this line in the spectrum of quasars does {\em not} prohibit denser material having a significant contribution to the spectrum.

% The model produces strong emission lines in \civ, \nv\ and \la,
% as well as a very weak \mg\ line.

The shapes and widths of these lines
match the composites fairly well, and the integrated 
line luminosities are significant when compared to observations.
However, there are a number of properties of the synthetic spectra
that are at odds with the observations. First, we find that the ratios of the 
equivalent width of the \la\ and \mg\ lines
to \civ\ is much lower than in the composite spectra. This problem 
has also been seen in simpler photoionization models for the BLR \citep{netzer1990}.
It may be that a larger region of very dense ($n_e\sim10^{10}$cm$^{-3}$) 
material is needed, this could correspond to a disc atmosphere or 
`transition region' (see e.g. Murray et al. 1995; Knigge et al. 1997). 
While modest changes to geometry may permit this, the initial grid search 
did not find a parameter space in which the \la\ or \mg\ equivalent widths
were significantly higher. Second, we find that equivalent widths increase
with inclinations, to the extent that significantly denser
models can match the line EWs fairly well at low inclinations but will then
possess overly strong red wings to the BAL P-Cygni profiles at high inclinations
(see also figure~\ref{fig:lobal}).
This angular dependence of EW is expected from a model in which the BLR radiates
more isotropically than the disc, which in this case is foreshortened and limb darkened.
It appears to contradict observations, which show remarkably uniform emission
line properties in quasars and BALQSOs \citep{weymann1991,dipompeo2012b}. 
The angular distribution of the disc 
continuum and line emission
is clearly crucially important in determining the emergent broad 
line equivalent widths, as suggested by, e.g., the analysis of \cite{risaliti2011}. 
We shall explore this question further in a future study.  
% However, the line-to-continuum ratios 
% at low inclinations in our model are significantly weaker than the quasar 
% composites. Increasing the density of the outflow, by altering the mass 
% loss rate  or velocity law, can produce more line emission.
% However, the red wing of the BAL profiles is generally stronger than 
% seen in BALQSO spectra and composites. This suggests a fairly 
% fundamental problem with a geometric unification model in which 
% The line equivalent widths at high inclinations are
% significantly affected by disc foreshortening and limb darkening.
% The angular distribution of the disc radiation is clearly
% crucially important in determining the emergent line ratios.  
% {\bf KSL: This is a bit oddly explained. Our problem is at low inclination.  Note that I am a bit concerned with Fig 3; the model spectra and the observations look very different.}

@@

\begin{figure*}
\centering
\includegraphics[width=0.8\textwidth]{figures/lx_a05_pre.png}
\caption
{
X-ray ($2$~keV) luminosity of the our clumped model (red squares) 
and the H13 model (blue squares), plotted against monochromatic luminosity 
at 2500\AA. The points are labeled according to inclination; angles
$>70^\circ$ correspond to BALs in our scheme (see figure 4).
Also plotted are masurements from 
the COMBI-7 AGN and the BQS samples (Steffen et al. 2006) and the Saez et al. (2012) 
sample of BALQSOs. The dotted line shows the best fit relation for non-BALQSOs 
from Steffen et al. (2006).
}
\label{fig:xray}
\end{figure*}

@@

The high inclination, `BALQSO-like' viewing angles show moderate agreement with the data,
and are X-ray weak due to bound-free and electron scattering opacities in the wind.
Typically, BALQSOs show strong X-ray absorption with columns 
of $N_H\sim10^{23}~\rm{cm^{-2}}$ 
\citep{green1996,mathur2000,green2001,grupemathur2003}.
This is often cited as evidence that the BAL outflow is shielded from
the X-ray source, especially as sources with strong X-ray absorption tend
to exhibit deep BAL troughs and high outflow velocities 
\citep{brandt2000,laorbrandt2002,gallagher2006}.
Our results imply that the clumpy BAL outflow
itself can be responsible for the strong X-ray absorption, 
and supports Hamann et al.'s (2013) suggestion that 
this explains the weaker X-ray absorption in mini-BALs 
compared to BALQSOs.

@@

\begin{figure*}
\centering
\includegraphics[width=0.8\textwidth]{figures/c4_angles.png}
\caption
{
\civ , \mg , \al\ and Fe~\textsc{ii} line profiles for viewing angles
from $72-79^\circ$. The profiles are plotted relative to the local
continuum with an offset applied for clarity. Lower ionization
profiles appear at a subset of high inclinations, compared
to the ubiquitous \civ\ profile.
}
\label{fig:lobal}
\end{figure*}

At high inclinations, the synthetic spectra exhibit blue-shifted BALs in \al\ and \mg --
the absorption lines seen in LoBALQSOs, and we even see absorption in Fe~\textsc{ii}
at the highest inclinations. Line profiles in velocity space 
for \civ, \al\ and \mg, are shown in figure~\ref{fig:lobal} for a range
of BALQSO viewing angles. We find that ionization stratification
of the wind causes lower ionization material to have a smaller covering factor, 
as demonstrated by figures~\ref{fig:wind} and \ref{fig:lobal}.
This confirms the behaviour expected from a unification model such as Elvis (2000). 
LoBALs are only present at viewing angles close to edge-on ($i>75^\circ$),
as predicted by polarisation results \citep{brotherton1997}.
As observed in a BALQSO sample by \cite{filizak2014}, we find that
BAL troughs are wider and deeper when low ionization absorption features are present,
and high ionization lines have higher blue-edge velocities than the 
low ionization species.
There is also a correlation between the strength of LoBAL features
and the amount of continuum attenuation at that sightline, particularly
blueward of the Lyman edge as the low ionization base 
intersects the line-of-sight. 
A model such as this therefore predicts that LoBALQSOs and FeLoBALQSOs 
have stronger Lyman edge absorption and 
are more Compton-thick than HiBALQSOs and Type 1 quasars.
An edge-on scenario also offers a potential explanation for the rarity of LoBAL and
FeLoBAL quasars, due to a foreshortened and attenuated continuum, 
although BAL fraction inferences are fraught with complex selection 
effects \citep{goodrich1997,krolikvoit1998}.

\subsection{Parameter Sensitivity}

\label{sec:param_sens}

\begin{figure*}
\centering
\includegraphics[width=1.0\textwidth]{figures/scatter_c4_grid.png}
\caption
{
The EW of the \civ~$1550$\AA\ line at $20^\circ$ plotted against a) the 
$BI$ of \civ~$1550$\AA\ at $75^\circ$, b) the EW of the \mg~$2800$\AA\ line 
at $20^\circ$ and c) the EW of \la\ at $20^\circ$. The circles correspond 
to the simulation grid for two different values of $f_V$, and the fiducial 
model is marked with an orange star. 
We also show a higher X-ray luminosity model and a higher mass loss rate
with a red star.
}
\label{fig:grid}
\end{figure*}

Having selected an individual fiducial model from the simulation grid, it is important
to briefly explore how specialised this model is, and how small parameter
changes can affect the synthetic spectra. Figure~\ref{fig:grid}
shows the EW at a low inclination, and $BI$ at a high inclination for the simulation
grid. A few conclusions can be drawn from this plot straight away. 
First, we find that almost all the models with $f_V=0.1$ are over-ionized, and 
fail to produce strong \civ\ BALs or emission lines. Second, we find that 
it is difficult to significantly increase line emission while
keeping the luminosity and mass loss rate of the system fixed. 
We show an additional point on figure 7 corresponding to a model with an order of
magnitude higher X-ray luminosity and double the mass loss rate. As expected, 
this results in far higher line equivalent widths, but fails to produce BALs because
the collisionally excited emission swamps the BAL profile. In addition,
this model would lie well above the expected $L_{2kev}-L_{2500}$ 
relation in figure 5. Such a high X-ray luminosity could therefore 
not be the cause of the strong line emission seen in {\em all} Type 1 quasars.

The parameter search presented here is by no means exhaustive, and
we may be limited by the specific parameterisation (from SV93) of the outflow 
kinematics we have used. Nevertheless, we suggest that the angular distribution
of {\em both} the line and continuum emission is perhaps the crucial 
aspect to understand.
With this in mind, obtaining reliable orientation indicators appears
to be a crucial observational task if we are to
further our understanding of BAL outflows 
and their connection, or lack thereof, to the broad line region.

@@

%  \section{Discussion: Parameter Sensitivity and Inclination Dependence}

%  We have demonstrated that our fiducial model successfully reproduces much of
%  the phenomenology of luminous AGN and quasars, with a few exceptions.
%  Here, we discuss these shortcomings, by discussing parameter sensitivity
%  and the dependence of line EWs on inclination.

% \subsection{The Angular Distribution of Line And Continuum Emission}
% \label{sec:angular}

% \begin{figure}
% \centering
% \includegraphics[width=0.49\textwidth]{figures/hist_ew.png}
% \caption
% {
% {\sl Top Panel:} A histogram of $1/2$ equivalent widths ($W_{\lambda}/2$)
% for BALQSOs and non-BAL quasars from the {\sl SDSS DR7} quasar sample (Shen et al. 2011).
% {\sl Bottom Panel:} $W_{\lambda,RW}$ as a function of inclination in our example model (black line).
% The green shaded area indicates viewing angles which look into the wind (`BALQSO-like').
% }
% \label{fig:ew}
% \end{figure}

% % \begin{figure}
% % \centering
% % \includegraphics[width=0.45\textwidth]{figures/ew.png}
% % \caption
% % {
% % $W_{\lambda,RW}$ as a function of inclination in our example model (red line).
% % We also the mean $1/2$ equivalent width from DiPompeo et al. (2012) as a horizontal
% % line, with the shaded area corresponding to the standard deviation.
% % }
% % \label{fig:ew}
% % \end{figure}

% In order to quantitatively assess how emission lines change with 
% inclination when blue-shifted absorption 
% may affect the line profile, we define the `red wing equivalent width' ($W_{\lambda,RW}$) as
% \begin{equation}
% W_{\lambda,RW} = \int_{\lambda_0}^{\lambda^\prime} \left( 1 - \frac{F_\lambda}{F_0} \right) d\lambda
% \label{rwew}
% \end{equation}
% where $F_0$ is the continuum flux and the integral is calculated from $\lambda_0$, line centre,
% to a wavelength $\lambda^\prime$ where the flux has returned to the continuum level.
% This quantity is shown as a function of inclination in figure 5 for the \civ\ and \mg\ UV lines.
% We also show the $W_{\lambda,RW}$ expected from isotropic line emission and a foreshortened 
% and limb darkened disc  as well as $1/2$ equivalent widths from \cite{dipompeo2012b}.
